---
title: "p8105_hw2_ym3197"
author: "Yichaun Ma"
date: "2026-10-06"
output: github_document
---

```{r}
library(tidyverse)
```
problem1
```{r}

nyc_transit <- read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", na = c("NA", "", "."))



dim(nyc_transit)


nyc_clean <- nyc_transit |> 
  rename_with(~ str_replace_all(tolower(.), " ", "_")) |> 
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    starts_with("route"), 
    entrance_type,
    entry,
    vending,
    ada
  ) |> 
  mutate(
    
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE,
      .default = NA
    ),
 
    across(starts_with("route"), as.character)
  )



dim(nyc_clean)
print("Entry variable conversion check:")
table(nyc_clean$entry, useNA = "ifany")
```

```{r}

distinct_stations <- nyc_clean |> 
  distinct(station_name, line) |> 
  nrow()
print(paste("1. Number of distinct stations:", distinct_stations))

ada_stations <- nyc_clean |> 
  filter(ada == TRUE) |> 
  distinct(station_name, line) |> 
  nrow()
print(paste("2. Number of ADA compliant stations:", ada_stations))


total_no_vending <- nyc_clean |> 
  filter(vending == "NO") |> 
  nrow()
print(paste("   Total entrances without vending (denominator):", total_no_vending))

no_vending_entry <- nyc_clean |> 
  filter(vending == "NO", entry == TRUE) |> 
  nrow()
print(paste("   Entrances without vending that allow entry (numerator):", no_vending_entry))

proportion_no_vending <- no_vending_entry / total_no_vending
print(paste("3. Proportion:", round(proportion_no_vending, 4)))

```

```{r}

nyc_long <- nyc_clean |> 
  pivot_longer(
    cols = starts_with("route"),
    names_to = "route_number",
    values_to = "route",
    values_drop_na = TRUE
  ) |> 
  mutate(route = str_trim(route))



dim(nyc_long)


a_train_stations <- nyc_long |> 
  filter(route == "A") |> 
  distinct(station_name, line) |> 
  nrow()
print(paste("4. Number of distinct stations served by A train:", a_train_stations))


a_train_ada <- nyc_long |> 
  filter(route == "A", ada == TRUE) |> 
  distinct(station_name, line) |> 
  nrow()
print(paste("5. A train stations that are ADA compliant:", a_train_ada))
```
problem 2

read and clean data
```{r}
# Problem 2


mr_trash_wheel =
  readxl::read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N1000"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.numeric(year),
    month = str_trim(month),
    sports_balls = as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel"
  )

```

read and clean data
```{r}
professor_trash_wheel =
  readxl::read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M1000"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.numeric(year),
    month = str_trim(month),
    trash_wheel = "Professor Trash Wheel"
  )

```



read and clean data
```{r}
gwynnda =
  readxl::read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynnda the Good Wheel of the W",
    range = "A2:L1000"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.numeric(year),
    month = str_trim(month),
    trash_wheel = "Gwynnda"
  )
```



```{r}
mr_trash_wheel =
  mr_trash_wheel |>
  mutate(
    across(
      any_of(c(
        "year",
        "weight_tons",
        "volume_cubic_yards",
        "plastic_bottles",
        "polystyrene",
        "cigarette_butts",
        "plastic_bags",
        "wrappers",
        "homes_powered"
      )),
      as.numeric
    )
  )

professor_trash_wheel =
  professor_trash_wheel |>
  mutate(
    across(
      any_of(c(
        "year",
        "weight_tons",
        "volume_cubic_yards",
        "plastic_bottles",
        "polystyrene",
        "cigarette_butts",
        "plastic_bags",
        "wrappers",
        "homes_powered"
      )),
      as.numeric
    )
  )

gwynnda =
  gwynnda |>
  mutate(
    across(
      any_of(c(
        "year",
        "weight_tons",
        "volume_cubic_yards",
        "plastic_bottles",
        "polystyrene",
        "cigarette_butts",
        "plastic_bags",
        "wrappers",
        "homes_powered"
      )),
      as.numeric
    )
  )
```


combined three data
```{r}

trash_wheel_df =
  bind_rows(
    mr_trash_wheel,
    professor_trash_wheel,
    gwynnda
  )

nrow(trash_wheel_df)

trash_wheel_df =
  bind_rows(
    mr_trash_wheel,
    professor_trash_wheel,
    gwynnda
  )


trash_wheel_df

nrow(trash_wheel_df)

names(trash_wheel_df)

professor_trash_wheel |>
  summarize(
    total_weight = sum(weight_tons, na.rm = TRUE)
  )
```
 

total cigaretee butts collected by Gwynnda in June 2022
```{r}
gwynnda |>
  filter(
    year == 2022,
    month == "June"
  ) |>
  summarize(
    total_cigarettes = sum(cigarette_butts, na.rm = TRUE)
  )




```

Problem 3

Import and clean baseline demographics
```{r}

mci_baseline =
  read_csv(
    "data/data_mci/MCI_baseline.csv",
    skip = 1,
    na = c("NA", ".")
  ) |>
  janitor::clean_names() |>
  filter(
    is.na(age_at_onset) | age_at_onset > current_age
  ) |>
  mutate(
    sex = case_match(
      sex,
      0 ~ "female",
      1 ~ "male"
    ),
    apoe4 = case_match(
      apoe4,
      0 ~ "non-carrier",
      1 ~ "carrier"
    )
  )

```

Number of participants recruited
```{r}
nrow(mci_baseline)

```
Number of participants who developed MCI
```{r}


mci_baseline |>
  filter(!is.na(age_at_onset)) |>
  nrow()
```

Average baseline age
```{r}

mci_baseline |>
  summarize(
    mean_age = mean(current_age, na.rm = TRUE)
  )
```

Proportion of female APOE4 carriers
```{r}

mci_baseline |>
  filter(sex == "female") |>
  summarize(
    female_apoe4_proportion = mean(apoe4 == "carrier")
  )
```

this is import and clean amyloid data
```{r}


mci_amyloid =
  read_csv(
    "data/data_mci/mci_amyloid.csv",
    skip = 1,
    na = c("NA", "")
  ) |>
  janitor::clean_names() |>
  rename(
    id = study_id,
    time_0 = baseline
  ) |>
  pivot_longer(
    time_0:time_8,
    names_to = "time",
    values_to = "amyloid"
  ) |>
  mutate(
    time = case_match(
      time,
      "time_0" ~ 0,
      "time_2" ~ 2,
      "time_4" ~ 4,
      "time_6" ~ 6,
      "time_8" ~ 8
    )
  )
```

Number of participants in amyloid dataset
```{r}

mci_amyloid |>
  distinct(id) |>
  nrow()
```
Participants only in baseline dataset
```{r}

baseline_only =
  anti_join(
    mci_baseline,
    mci_amyloid,
    by = "id"
  )

nrow(baseline_only)
```

Participants only in amyloid dataset
```{r}


amyloid_only =
  anti_join(
    mci_amyloid,
    mci_baseline,
    by = "id"
  )

amyloid_only |>
  distinct(id) |>
  nrow()
```

Number of amyloid-only observations
```{r}

nrow(amyloid_only)
```
Combine baseline and amyloid datasets
```{r}

mci_complete =
  inner_join(
    mci_baseline,
    mci_amyloid,
    by = "id"
  )

```
Number of participants in combined dataset
```{r}

mci_complete |>
  distinct(id) |>
  nrow()
```

Number of observations in combined dataset
```{r}
nrow(mci_complete)
```

View combined dataset
```{r}

mci_complete

names(mci_complete)
```



Export combined dataset
```{r}

write_csv(
  mci_complete,
  "data/data_mci/mci_complete.csv"
)
```




```{r}

```



```{r}

```



```{r}

```


```{r}

```



```{r}

```